### 2021-02-15 Thesis/article.tex
Initial commit

\documentclass[a4paper]{article}
\usepackage[margin=3cm]{geometry}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{lipsum}
\newcommand{\todo}[1]{\textbf{\textcolor{red}{#1}}}
\title{Title}
\author{Johann Dorn}

\begin{document}
\maketitle
\iffalse
my notes
\fi

\section{Introduction}
scarce data may lead to overfitting\cite{Lecun1995conv}
\subsection{RAMAN}
\clearpage
\end{document}