### 2021-02-15 Thesis/article.tex
thesis mail to theo about sol-gel process

\documentclass[a4paper]{article}
\usepackage[margin=3cm]{geometry}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{lipsum}
\newcommand{\todo}[1]{\textbf{\textcolor{red}{#1}}}
\title{Title}
\author{Johann Dorn}

\begin{document}
\maketitle
\iffalse
my notes
\fi

Making of the solution for the sol-gel process:
For a single concentrated solution 0.05 ml of Zr(IV)Propoxide are added while stirring to 4.95 ml of Buthan-1-ol and stirred for 15 minutes. 
0.013 ml (or one molar equvilent of Zr) of Acetylacetone is added to the stirring solution. After another 15 minutes 1 ml of Aceteicacid is added and stirred for 30 minutes to stabilize the solution up to 24h.

The concentration can be increased up to 5 times being stable for a minimum of 4 hours. The sol-gel process produces am homogeneous transparent crystalline zirconia oxide layer. homogeneity can be mainly controlled via blade velocity and temperature and layers can be stacked.

\section{Introduction}
scarce data may lead to overfitting\cite{Lecun1995conv}
\subsection{RAMAN}
\clearpage
\end{document}