### 2021-12-12 Thesis/chapter/aims-and-objectives.tex
thesis split into chapters

The aim of this work is to develop a non-conducting layer on steel based on  non-toxic materials like aluminium or zirconium via doctor blading. 
This layer can then be used as insulator for CIGS modules on steel sheets.
Doctor blading - or tape casting - is a widely used precision coating method to apply thin films on large area surfaces\cite{Berni2004}.
%The method of doctor blading, which is in principle a sol-gel process, 
This sol-gel process was chosen over spin coating because of the availability to the 
industrial partner.
In order to optimize the resulting layer the multitude of parameters was optimized with an particle swarm optimization ansatz. %machine learning 
The conductivity (dependent variable), the number of layers and calcination time (both independent variables) should be minimized.
%The conductivity should be as small as possible.
%The number of layers should be held small and the calcination heating rate should be maximized. 
\iffalse
Defensio: Argumente in der Arbeit noch mal sichten, Feedback von Betreuungsperson im Kopf haben - da könnten Fragen kommen, 
sich selbst aufnehmen

Präsentieren: Visualisierungen sinnvoll? 
Antworten überlegen/Argumente überlegen
Limitation, Rahmen der Arbeit
Begründung für die eigene Vorgehensweise
\fi