### 2022-02-28 Thesis/chapter/07-outlook.tex

Making of the solution for the sol-gel process:
For a single concentrated solution \ml{0.05} of \gls{zrpro} are added while stirring to \ml{4.95} of \gls{buoh} and stirred for \minutes{15}. 
\ml{0.013} (or one molar equvilent of Zr) of \gls{acac} is added to the stirring solution. 
After another \minutes{15} \ml{1} of acetic acid is added and stirred for \minutes{30} to stabilize the solution up to \h{24}.

The concentration can be increased up to 5 times being stable for a minimum of \h{4}. 
The sol-gel process produces am homogeneous transparent crystalline zirconia oxide layer. 
homogeneity can be mainly controlled via blade velocity and temperature and layers can be stacked.

It should have been alos verglichen with grid search with comparable size
but most time was used to find a vernuenfig base recipe and process

It is still very human 
Der process is - as it the case with all ML and most fitting processes - is very abhaengig von hyper parameters, 
In the current work population size, number of generations, and most importantly boundaries (grenzen).