### 2022-10-16 Thesis/chapter/02-aims-and-objectives.tex
free writing einarbeiten

The aim of this work is to develop a non-conducting layer on steel based on  non-toxic materials like aluminium or zirconium via doctor blading. 
This layer can then be used as insulator for CIGS modules on steel sheets.
Doctor blading - or tape casting - is a widely used precision coating method to apply thin films on large area surfaces\cite{Berni2004}.
%The method of doctor blading, which is in principle a sol-gel process, 
This sol-gel process was chosen over spin coating because of 
its scalability.
%the availability to the industrial partner.
In order to optimize the resulting layer the multitude of parameters was optimized with an particle swarm optimization ansatz. %machine learning 
The conductivity (dependent variable), the number of layers and calcination time (both independent variables) should be minimized.
%The conductivity should be as small as possible.
%The number of layers should be held small and the calcination heating rate should be maximized. 
\iffalse
Defensio: Argumente in der Arbeit noch mal sichten, Feedback von Betreuungsperson im Kopf haben - da könnten Fragen kommen, 
sich selbst aufnehmen

Präsentieren: Visualisierungen sinnvoll? 
Antworten überlegen/Argumente überlegen
Limitation, Rahmen der Arbeit
Begründung für die eigene Vorgehensweise
\fi
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{2022-06-19-Free-Writing}
The problem which presents it self is in principle a search problem.
A large space of possible combinations of experimental anordnungen presents it self
and out of these we want to find the optimial parameter combination. 
The best way with unlimitied resources would be the brute-force approach to try every possibility. 
But due to limited time this is unfeasable. 
But a new questions arises: how should the ausmasse of the search space be eingeschraenkt? 
What are the upper and lower limits and which intervals/step size/grid size should be used. 
Also even before, we have to answer which dimensions/parameters should we include in the search? 
Which parameters should we keep constant and at which value? 
And finally which features should we use to measure the goodness of an experiment?

I tried different methods to answer the first X questions: Latin square and Plackett-Burman design.
Unfortunately, I was faced with similar questions. How many levels and which end points? 
latin squares are about orthogonality 
PB is a factorial experiment desing in contrast to one-fator-at-a-time methods.
reliability and validity are low due to experiments not replicated

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%