### 2022-10-21 Thesis/chapter/06-results-and-discussion.tex
start results emma

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
and cite something pro forma \cite{ncbi1butanol}
%\input{chapter/060-results-material.tex}
\input{chapter/061-results-emma.tex}
%\input{chapter/062-discussion.tex}