### 2023-02-10 Thesis/chapter/032-alt-theory-calculations.tex
svm chaos

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Machine Learning and Statistics}

\subsubsection{Introduction to AI, ML and statistics}
\textbf{What is AI?} It is the interdisciplinary field of making a artificial inteligent agent. 
What is intelligence though?
Intelligence is not easily defined. Some definitions I agree with is that intelligence 
includes learning from experience, finding solutions to new problems. 
Creativity is also substencial for intelligence. 
It is a goal in the field AI to immitate (and surpass) human intelligence. 
The turing tests aims to test a computers intelligience by testing if it can trick a person into believing that the computer is person while a person tries to make the person believe that they are in fact a person. 
Approaches range from knowledge based systems over hard coded intstructions to machine learning. 
%
\textbf{Machine Learning} is the art of telling a machine or computer how to gain knowlegde from data (aka experience). 
There are different problems and solutions. 
The problems include image recognition, voice recognition, natural language processing, data prediction, classification, recommendation systems and so on. 
%
In order to solve these problems machine learning resorts to statistical methods such as
linear regression, ridge regression. 
Further machine learning methods not regularly seen in statistics are the 
kernel trick, support vector machines, genetic algorithms and particle swarm algorithms. 
The last two are special in the sense that they don't provide a certain regression method,
but rather a frame work to choose the next sample which should be tested in a optimisation. 
\td{mention classification vs regression and supervised vs supervised}
%
\textbf{Statistics} is a subfield of mathematics and emerged from the aim to summerzie large data sets 
and extract information. In contrast to descriptive statistic, there is stochastic. 
Stochastic statistic aims on predicting data points with probabilities. 
%
\textbf{Linear regression} is one of the simplest methods to predict data. 
It besticht by its computational simplicity and easy interpretation. 
%Let's take this equation
\begin{equation}
\mathbf{y} = \mathbf{X} \mathbf{k} +c 
\end{equation}
Where $\mathbf{y}$ (size $\mathbb{R}^n$, where $n$ is the number of data points) is a vector of dependent variables and $\mathbf{X}$ (size $\mathbb{R}^{n\times d}$, where $d$ is the the number of independent variables) is a matrix of independent variables. 
The parameters $\mathbf{k}$ (size $\mathbb{R}^d$) and $c$ (size $\mathbb{R}$) are chosen by minimizing an objective function (loss function in \gls{ml} chargon).
A typical function to minimize are the L1 and L2 error, i.e. \gls{mae} and \gls{mse}, respectively.
\begin{align}
L_1&= \sum |y_i - \hat y_i| = \sum \sum |y_i - x_{i,j}\cdot k_i+c| \\
L_2&= \sum(y_i - \hat y_i)^2 = \sum (y_i - x_{i,j}\cdot k_i+c)^2
\end{align}
%

\textbf{Ridge Regression} is like linear regression but with an extra term which penalizes steep regression functions.
This extra term is scaled by a correction parameter $\lambda$. 
The larger $\lambda$ the larger regularisation and the flatter the regression functions. 
If $\lambda$ were infinite, then the regression function would be a flat line. 
The correction term avoids overfitting. 
\begin{equation}
    L_{RR} = L_2 + \lambda \sum k_i = \sum(y_i - \hat y_i)^2 + \lambda \sum k_i
\end{equation}

\textbf{Kernel Ridge Regression}
\gls{krr} combines ridge regression with the kernel method. 
A kernel transforms data in such a way that a linear hyperplane (a line in one dimension, 
a plane in two dimensions) can fit data in regression problem or seperate the data in classification problem. 
A kernel is some kind of similarity measure. 
The following equations show definitions for 
linear (eq.~\ref{eq:lin-kernel}), 
polynomial (eq.~\ref{eq:pol-kernel}), 
sigmoidal (eq.~\ref{eq:sig-kernel}) and 
radial basis functionial kernel~(eq. \ref{eq:rbf-kernel}).
\begin{align}
    \label{eq:lin-kernel}
    k_{lin}(\mathbf{x},\mathbf{y}) &= \mathbf{x}^{\top} \mathbf{y} \\
    \label{eq:pol-kernel}
    k_{pol}(\mathbf{x}, \mathbf{y}) &= (\gamma \mathbf{x}^{\top} \mathbf{y} + c_0)^d \\
    \label{eq:sig-kernel}
    k_{sig}(\mathbf{x}, \mathbf{y}) &= \text{tanh}(\gamma \mathbf{x}^{\top} \mathbf{y} + c_0) \\
    \label{eq:rbf-kernel}
    k_{rbf}(\mathbf{x}, \mathbf{y}) &= \text{exp}(- \gamma \| \mathbf{x} -  \mathbf{y} \|^2) 
\end{align}
These kernels fulfill the requirements of nonnegativity, symmetry, linearity\cite{rupp2015machine}.
\td{Note that when $\lambda$ tends to zero, the regularized loss function becomes the OLS loss function.
When $\lambda$ tends to infinity, we get an intercept-only model}
\td{A critique of ridge regression is that all the variables tend to end up in the model. The model only shrinks the coefficients.}
% from https://medium.com/@zxr.nju/the-classical-linear-regression-model-is-good-why-do-we-need-regularization-c89dba10c8eb
%
% https://scikit-learn.org/stable/modules/metrics.html#polynomial-kernel
%

\textbf{Support vector machines} 
\Gls{svm} are versatile machine learning algorithm first mentioned in 1992\cite{boser1992training}. 
\Gls{svm} is mainly used for classification, but can also be used for regression. 
Classifiction works by spanning a hyperplane between two categories in a way that the closest points from each category have the largest distance to the hyperplane. 
The distance from the points to the hyperplane is called margin.
These points are called support vectors and are used to define the hyperplane 
A \gls{svm} avoids overfitting by only using a subset - the support vectors - to fit the model. 
\td{The goal is to fit $w$ and $b$ such that the margin (the distance from the descission boundary is maximal.
As this the margin is inversely proportional to $\|w\|$, we aim to minimize $\|w\|$.
With the contraint that all data points are correctly classified $y_i(x_iw-b)\geqslant1$.}
In the same manner as no data points are found within the margin for classification, 
no data points should be found outside of the margin. 
If the data is not seperable by a hyperplane the kernel trick can be used. 
The downside of \gls{svm} is that it is very
\td{$\text{min}\frac{1}{N} \sum_{i=1}^{N} max(0,1- y_i(\mathbf{x}x_i-b)) + \lambda \|w\|$
The downside of \gls{svm} is that it is very sensitive to outliers.}

$\min_{w,b} \|w\|$

$ y_i(w \cdot x_i - b)\leqslant 1$

Summary of SVM from Cherkassky's \textit{Learning from Data}\cite{cherkassky1998learning}
The \gls{svm} was initially developed by V. Vapnik for the binary classification of seperable data, then improved to handle non-sperable data and eventually adapted to solve regression problems.
The concepts of \gls{svm} will be discussed in the same chronological order.

The decision boundary can be expressed as an hyperplane
\begin{equation}
	\hat{\mathbf{y}} = D(\mathbf{x}) = \mathbf{w} \cdot \mathbf{x} + w_0 
\end{equation}
with $
\mathbf{y} \in \lbrace +1, -1 \rbrace^n ,\,
\mathbf{w} \in \mathbb{R}^d ,\,
\mathbf{x} \in \mathbb{R}^{n \times d} 
$.
A hyperplane which separate the data will satisfy the following equation: 
\begin{equation}
	\label{eq:svm01}
	y_i(\mathbf{w} \cdot \mathbf{x}_i + w_0 ) \geqslant 1, \qquad i=1,\dots,n
\end{equation}
The distance from the decision boundary to the nearest points shall be denoted as $\tau$ and should be maximized. 
Inequality \ref{eq:svm01} is an equation for the support vectors. 
If we now take a vector from a positive support vector $x_+$ to a negative support vector $x_-$ and project it on to the unit vector of $w$ we get the width of the margin. 
\begin{equation}
	\label{eq:svm02}
\begin{aligned} 
	\text{width} &= (x_+ - x_-) \cdot \frac{w}{\|w\|} \\
		&= ( (1-b)/w - (-1-b)/w ) ) \cdot \frac{w}{\|w\|} \\
		&= \frac{2}{w} \cdot \frac{w}{\|w\|} = \frac{2}{\|w\|}
\end{aligned}
\end{equation}
\td{draw sketch and make \ref{eq:svm02} on two lines}

but before are the hyperplanes which go through the support vectors for positive sample: 
\begin{equation}
	y
\end{equation}

\begin{itemize}
    \item \sout{intro to stat and ml}
    \item \sout{say that although different will be abgehandelt in the same chapter}
    \item \sout{basic statistics: MAE vs MSE, R, std, F-test} 
    \item \sout{lin reg}
    \item \sout{ridge regression}
    \item \sout{kernel ridge regression}
    \item \sout{support vector machine}
    \item anova (categorical), PCA, step wise regression
    \item all previous ananlyse data to predict, but don't tell me which data point is needed next 
    \item DOE (add data acquisition)
    \item genetic algorithm and pso (they combine data gen with regression/prediction and optimisation, actually incorporate prediction into selection process but don't per se specify which pred and optimisation)
    \item comment: always tradeoff between complexity and interpretability
    \item quesiton: which paragraph by GPT2?? 
\end{itemize}