### 2023-02-26 Thesis/chapter/07-outlook.tex
start outlook

\todo{freewriting:} 
what did i want to do? 
The goal of this work was to find a sol-gel process resulting in a insulating zirconium ceramic on top of steel.
This aufbau then can be used for thin layer photo voltaics. 
The ceramic needs to be homoegeneous and defect free. 
what was done? 
A suitable recipe was found and optimized via \gls{emma}. 
welche methoden wurden dafür verwendet? 
after the optimsation the data was analysed with further/different machine learning methods like kernel ridge regression, support vector regression and linear regression. 
was wurde herausgefunden? 
A strong correlation of calcination temperature and conductance in the bereich of 300--400\oc{}. 
was kann noch herausgefundne werden und wie?

was kann noch veraendert werden? 
humidity 
solution age
vdb and tdb on g and phd 
Tcal on g and phd

Making of the solution for the sol-gel process:
For a single concentrated solution \ml{0.05} of \gls{zrpro} are added while stirring to \ml{4.95} of \gls{buoh} and stirred for \minutes{15}. 
\ml{0.013} (or one molar equvilent of Zr) of \gls{acac} is added to the stirring solution. 
After another \minutes{15} \ml{1} of acetic acid is added and stirred for \minutes{30} to stabilize the solution up to \h{24}.

The concentration can be increased up to 5 times being stable for a minimum of \h{4}. 
The sol-gel process produces am homogeneous transparent crystalline zirconia oxide layer. 
homogeneity can be mainly controlled via blade velocity and temperature and layers can be stacked.

It should have been also verglichen with grid search with comparable size
but most time was used to find a vernuenfig base recipe and process

It is still very human 
Der process is - as it the case with all ML and most fitting processes - is very abhaengig von hyper parameters, 
In the current work population size, number of generations, and most importantly boundaries (grenzen).